---
title: "Italy_Analysis"
author: "Afia Amoako"
date: '2020-03-29'
output:
  html_document: default
  pdf_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r include=FALSE, results='hide'}
#required package
library(httr)
library(tidyverse)
library(lubridate)
library(knitr)
```

##Introduction

Analysis for the impact of intervention on COVID transmission is adapted from Wang, H., Wang, Z., Dong, Y., Chang, R., Xu, C., Yu, X., Zhang, S., Tsamlag, L., Shang, M., Huang, J., Wang, Y., Xu, G., Shen, T., Zhang, X., & Cai, Y. (2020). Phase-adjusted estimation of the number of Coronavirus Disease 2019 cases in Wuhan, China. Cell Discovery, 6(1), 1–8. https://doi.org/10.1038/s41421-020-0148-0


###Importing Data 
```{r results='hide'}
COVID <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
head(COVID, 10)
summary(COVID)

#things to do, downloads deaths, add death to all models, plot deaths 
```

###Cleaning Italy Table 
```{r results='hide', fig.show='hide'}
#CASES
#cleaning removing province, longititude and latitude 
COVID.clean <- COVID[ ,-c(1,3,4)]

#subset Italy cumulative cases
COVID.italy <- COVID.clean %>% filter(Country.Region == "Italy")
#creating cumulative cases table for  Italy 
start_date <- mdy("1-22-2020")
end_date <- mdy("4-7-2020")
intervals <- seq(from = start_date, to=end_date, by = "day")
cases <- as.numeric(COVID.italy[1, -c(1, 79:81)])
italy.cases <- as.data.frame(cbind(Date = intervals, Cases = cases))
italy.cases$Date <- as.Date(italy.cases$Date, origin ="1970-01-01")

#plot.cummulative.cases
plot(italy.cases$Date, type = "l", lwd = "3", col = "navy", italy.cases$Cases, main = "Cumulative Cases in Italy", xlab = "Date", ylab = "Cases")
```

###Model Validation: Estimating the number of cases based on different R0 values

Callibrating the high transmission based on current information (priors) to callibrate for first 40 days so basically when cases in Italy were growing quickly with little to no interventions
```{r results='hide' , fig.show='hide'}
#assumption: random mixing, person to person transmission, no difference in individual immunity, no deaths

# Parameter values
#initial variables per year
N_italy <- 60480000 #population of Italy
I_italy <- 2 #ikmported cases from Italy
E_italy <- 20 * I_italy #number potentially exposed
R_italy <- 0 
S_italy <- N_italy-E_italy-I_italy-R_italy
alpha <- 1/5.6 #incubation period
gamma <- 1/5 #recovery period
duration <- 5 #days 

# R0 and beta for different phases based on the impact of intervention to minimize transmission
beta_1_a <- 3.1/(N_italy*duration)
beta_2_a <- 3.5/(N_italy*duration)

#time specification
dt <- 1
start <- 0
end_italy <- 40
time_italy <- seq(start, end_italy, by = dt)
niter_italy <- (end_italy - start) / dt + 1
s_date_italy <- mdy("1-31-2020") #first cases reported in Rome
e_date_italy <- mdy("3-11-2020") #last up to date results
axis_italy <- seq(from = s_date_italy, to = e_date_italy, by = "day")


#create table
l <- 1 #index for beta_1: R0: 3.1
m <- 2 #index for beta_2: R0: 3.5

X_italy <- array(data = 0, dim = c(niter_italy, 5, 2)) #created an array to catch the results
colnames(X_italy) <- c("S", "E", "I","R", "New Case")
#3.1
X_italy[1, 1, l] <- S_italy #initial susuceptible population R0: 1.9
X_italy[1, 2, l] <- E_italy #initial latent population R0: 1.9
X_italy[1, 3, l] <- I_italy #initial infectious population R0: 1.9
X_italy[1, 4, l] <- R_italy #initial recovered population R0: 1.9
X_italy[1, 5, l] <- I_italy #initial infectious population R0: 1.9: using this row to count the number of new cases for cumulative incidence calculation.

#3.5
X_italy[1, 1, m] <- S_italy #initial susuceptible population R0: 2.6
X_italy[1, 2, m] <- E_italy #initial latent population R0: 2.6
X_italy[1, 3, m] <- I_italy #initial infectious population R0: 2.6
X_italy[1, 4, m] <- R_italy #initial recovered population R0: 2.6
X_italy[1, 5, m] <- I_italy #initial infectious population R0: 2.6: using this row to count the number of new cases for cumulative incidence calculation.

for (i in 2:niter_italy) { 
  #3.1
  X_italy[i,1, l] <- X_italy[i-1, 1, l] + dt * (-beta_1_a * X_italy[i-1,1,l] * (X_italy[i-1,3,l]) )  #S
  X_italy[i,2, l]<- X_italy[i-1,2, l] + dt * (beta_1_a * X_italy[i-1,1,l] *  (X_italy[i-1,3,l]) - alpha * X_italy[i-1,2,l]) #E
  X_italy[i,3, l] <- X_italy[i-1,3, l] + dt * (alpha * X_italy[i-1,2, l] - gamma * X_italy[i-1,3, l])  #I 
  X_italy[i,4, l] <- X_italy[i-1,4, l] + dt * (gamma * X_italy[i-1,3, l]) #R
  X_italy[i,5, l] <- dt * (alpha * X_italy[i-1,2, l]) #new cases added 
  
  #3.5
  X_italy[i,1, m] <- X_italy[i-1, 1, m] + dt * (-beta_2_a * X_italy[i-1,1,m] * (X_italy[i-1,3,m])  ) #S
  X_italy[i,2, m]<- X_italy[i-1,2, m] + dt * (beta_2_a * X_italy[i-1,1,m] * (X_italy[i-1,3,m]) - alpha * X_italy[i-1,2,m]) #E
  X_italy[i,3, m] <- X_italy[i-1,3, m] + dt * (alpha * X_italy[i-1,2, m] - gamma * X_italy[i-1,3, m])  #I 
  X_italy[i,4, m] <- X_italy[i-1,4, m] + dt * (gamma * X_italy[i-1,3, m]) #R    
  X_italy[i,5, m] <- dt * (alpha * X_italy[i-1,2, m]) #new cases added
   
}

first_beta_italy <- as.data.frame(X_italy[ , ,l])
second_beta_italy <- as.data.frame(X_italy[ , ,m])
first_beta_italy$Date <- axis_italy
second_beta_italy$Date <- axis_italy


#"Comparing Modeled Cumulative Cases in Italy with varying R0 values to current cases",
plot(italy.cases[c(10:50), 1], italy.cases[c(10:50), 2], type = 'l', xlab = "Date", ylab = "Cumulative Cases", col = "chartreuse4", lwd = 3, main = "Model Validation of \n varying R0 values ")
 lines(cumsum(first_beta_italy$`New Case`) ~ axis_italy, col = "pink", lwd = 3)
 lines(cumsum(second_beta_italy$`New Case`) ~ axis_italy, col = 'red', lwd = 3)
  legend("topleft", legend = c("Cumulative Cases \n in Italy", "Model R: 3.1","Model R:3.5"), fill = c("chartreuse4", "pink",  "red"), col = c("chartreuse4", "pink",  "red"), cex = 0.85, box.lty = 0)

```

###Phase Adjustment

####Current Dates
I applied interventions put in place to slow transmission looked at how that will change cases. I compared my values to the cumulative cases, looking at the current situation, my model approximation of the current situation and if the intervention actually creates an impact

```{r  results='hide' , fig.show='hide'}

##############################Intervention###########################################

#assumption: random mixing, person to person transmission, no difference in individual immunity, no deaths
# Parameter values
N_italy <- 60480000
I_italy <- 2 
E_italy <- 20 * I_italy
R_italy <- 0 
S_italy <- N_italy-E_italy-I_italy-R_italy
alpha <- 1/5.6 #incubation period
gamma <- 1/5 #recovery period
duration <-5 #days 
# R0 and beta for different phases based on the impact of intervention to minimize transmission #using infromation from more recent papers. Since R0 = ce/D, these interventions are targeting the ce (contact rate) to reduce transmision. 


R0_1 <- 1.5 #low end (the CI from JHU and authors)
R0_2 <- 2.5 #what was calculated as gthe potential R
R0_3 <- 3.5 #high end (the CI from JHU and authors)
R0_4 <- 0.5 #updated: reduces contact by up to 85% because of lockdown
beta_1 <- R0_1/(N_italy * duration)
beta_2 <- R0_2/(N_italy * duration)
beta_3 <- R0_3/(N_italy * duration)
beta_4 <- R0_4/(N_italy * duration)

#time specification
dt <- 1
start <- 0
end_italy_p <- 151 #days
time_italy_p <- seq(start, end_italy_p, by = dt)
niter_italy_p <- (end_italy_p - start) / dt + 1
s_date_italy_p <- mdy("1-31-2020")
e_date_italy_p <- mdy("6-30-2020")
axis_italy_p <- seq(from = s_date_italy_p, to = e_date_italy_p, by = "day")


X_italy_p <- array(data = 0, dim = c(niter_italy_p, 5)) #created an array to catch the results
colnames(X_italy_p) <- c("S", "E", "I","R", "New Cases")

X_italy_p[1, 1] <- S_italy #initial susuceptible population 
X_italy_p[1, 2] <- E_italy #initial latent population 
X_italy_p[1, 3] <- I_italy #initial infectious population 
X_italy_p[1, 4] <- R_italy #initial recovered population 
X_italy_p[1, 5] <- I_italy #new cases

for (i in 2:niter_italy_p) { 
  if(i < 25) {
  #phase 1 transmission with an R0 of 3.1 (January 31 - February 23rd)
  X_italy_p[i,1] <- X_italy_p[i-1, 1] + dt * (-beta_3 * X_italy_p[i-1,1] * X_italy_p[i-1,3] )  #S
  X_italy_p[i,2]<- X_italy_p[i-1,2] + dt * ((beta_3 * X_italy_p[i-1,1] * X_italy_p[i-1,3]) - alpha * X_italy_p[i-1,2]) #E
  X_italy_p[i,3] <- X_italy_p[i-1,3] + dt * (alpha * X_italy_p[i-1,2] - gamma * X_italy_p[i-1,3])  #I 
  X_italy_p[i,4] <- X_italy_p[i-1,4] + dt * (gamma * X_italy_p[i-1,3]) #R
  X_italy_p[i,5] <- dt * (alpha * X_italy_p[i-1,2])
  }
  else {
    if ( i >= 25 & i < 41) { 
  #phase 2 transmission with an R0 of 1.9 (February 24 -  March 8) blocking the north
  X_italy_p[i,1] <- X_italy_p[i-1, 1] + dt * (-beta_2 * X_italy_p[i-1,1] * X_italy_p[i-1,3] )  #S
  X_italy_p[i,2]<- X_italy_p[i-1,2] + dt * ((beta_2 * X_italy_p[i-1,1] * X_italy_p[i-1,3]) - alpha * X_italy_p[i-1,2]) #E
  X_italy_p[i,3] <- X_italy_p[i-1,3] + dt * (alpha * X_italy_p[i-1,2] - gamma * X_italy_p[i-1,3])  #I 
  X_italy_p[i,4] <- X_italy_p[i-1,4] + dt * (gamma * X_italy_p[i-1,3]) #R
  X_italy_p[i,5] <- dt * (alpha * X_italy_p[i-1,2])
    }
    else{
      if( i >= 41 & i < 55){
        #phase 3 transmission with an R0 of 0.65 (March 8 - March 10) North total Lockdown
        X_italy_p[i,1] <- X_italy_p[i-1, 1] + dt * (-beta_1 * X_italy_p[i-1,1] * X_italy_p[i-1,3] )  #S
        X_italy_p[i,2]<- X_italy_p[i-1,2] + dt * ((beta_1 * X_italy_p[i-1,1] * X_italy_p[i-1,3]) - alpha * X_italy_p[i-1,2]) #E
         X_italy_p[i,3] <- X_italy_p[i-1,3] + dt * (alpha * X_italy_p[i-1,2] - gamma * X_italy_p[i-1,3])  #I 
        X_italy_p[i,4] <- X_italy_p[i-1,4] + dt * (gamma * X_italy_p[i-1,3]) #R
         X_italy_p[i,5] <- dt * (alpha * X_italy_p[i-1,2])  
      }
      else{
        X_italy_p[i,1] <- X_italy_p[i-1, 1] + dt * (-beta_4 * X_italy_p[i-1,1] * X_italy_p[i-1,3])  #S
        X_italy_p[i,2]<- X_italy_p[i-1,2] + dt * ((beta_4 * X_italy_p[i-1,1] * X_italy_p[i-1,3]) - alpha * X_italy_p[i-1,2]) #E
         X_italy_p[i,3] <- X_italy_p[i-1,3] + dt * (alpha * X_italy_p[i-1,2] - gamma * X_italy_p[i-1,3])  #I 
        X_italy_p[i,4] <- X_italy_p[i-1,4] + dt * (gamma * X_italy_p[i-1,3]) #R
         X_italy_p[i,5] <- dt * (alpha * X_italy_p[i-1,2])  
      }
      }
  }
  }

X_italy_p <- as.data.frame(X_italy_p)
cumsum(X_italy_p$`New Cases`)
X_italy_p$date <- axis_italy_p

  
##############################Comparison with no iuntervention###########################################
#no intervention using high transmission rate 

# Parameter values
N_italy_n <- 60480000
I_italy_n <- 2 
E_italy_n <- 20 * I_italy
R_italy_n <- 0 
S_italy_n <- N_italy_n-E_italy_n-I_italy_n-R_italy_n
alpha <- 1/5.6 #incubation period
gamma <- 1/5 #recovery period
duration <-5 #days 

R0_3 <- 3.5 #high end (the CI from JHU and authors)
beta_3 <- R0_3/(N_italy * duration)

#time specification
dt <- 1
start <- 0
end_italy_p_n <- 151 #days
time_italy_p_n <- seq(start, end_italy_p_n, by = dt)
niter_italy_p_n <- (end_italy_p_n - start) / dt + 1
s_date_italy_p_n <- mdy("1-31-2020")
e_date_italy_p_n <- mdy("6-30-2020")
axis_italy_p_n <- seq(from = s_date_italy_p_n, to = e_date_italy_p_n, by = "day")

X_italy_p_n <- array(data = 0, dim = c(niter_italy_p_n, 5)) #created an array to catch the results
colnames(X_italy_p_n) <- c("S", "E", "I","R", "New Cases")

X_italy_p_n[1, 1] <- S_italy_n #initial susuceptible population 
X_italy_p_n[1, 2] <- E_italy_n #initial latent population 
X_italy_p_n[1, 3] <- I_italy_n #initial infectious population 
X_italy_p_n[1, 4] <- R_italy_n #initial recovered population 
X_italy_p_n[1, 5] <- I_italy_n #new cases


for(i in 2:niter_italy_p_n) {
  X_italy_p_n [i,1] <- X_italy_p_n [i-1, 1] + dt * (-beta_3 * X_italy_p_n [i-1,1] * X_italy_p_n[i-1,3] )  #S
  X_italy_p_n [i,2]<- X_italy_p_n [i-1,2] + dt * ((beta_3 * X_italy_p_n [i-1,1] * X_italy_p_n[i-1,3]) - alpha * X_italy_p_n[i-1,2]) #E
  X_italy_p_n [i,3] <- X_italy_p_n [i-1,3] + dt * (alpha * X_italy_p_n [i-1,2] - gamma * X_italy_p_n [i-1,3])  #I 
  X_italy_p_n [i,4] <- X_italy_p_n [i-1,4] + dt * (gamma * X_italy_p_n [i-1,3]) #R
  X_italy_p_n [i,5] <- dt * (alpha * X_italy_p_n [i-1,2]) #new cases
  }

X_italy_p_n <- as.data.frame(X_italy_p_n)
cumsum(X_italy_p_n$`New Cases`)
X_italy_p_n$date <- axis_italy_p
  

###########plots#################

#plot: number of new cases main = "Impact of Intervention on the Number of Infectious People Each Day",
plot(X_italy_p[c(1:24), 3] ~ axis_italy_p[c(1:24)] ,  col = 'steelblue4',  lwd = 3,  type = 'l', ylab = "Number of People Infectious", xlab = "Time (Days)",  xlim = as.Date(c("2020-01-31", "2020-06-30")), ylim = c(0,10000), main = "Number of People Infectious \n with Interventions")
  lines(X_italy_p[c(24:41), 3] ~ axis_italy_p[c(24:41)] , col = 'violetred4', lwd = 3)
  lines(X_italy_p[c(41:55), 3] ~ axis_italy_p[c(41:55)], col = 'gold', lwd = 3)
  lines(X_italy_p[c(55:152), 3] ~ axis_italy_p[c(55:152)], col = 'red', lwd = 3)
  lines(X_italy_p_n[c(1:152), 3] ~ axis_italy_p[c(1:152)] , col = 'navy', lwd = 3) #if we are assuming that transmission continued at a high rate
legend("topright", legend = c("R=3.5", "R=2.6", "R=1.5", "R=0.5", "No Intervention"), fill = c("steelblue4", "violetred4", "gold", "red", "navy"), col = c("steelblue4", "violetred4", "gold", "red", "navy"), cex = 0.75, box.lty = 0)

#cumulative cases comparison
#without intervention
plot(cumsum(X_italy_p_n[c(1:152), 5]) ~ axis_italy_p[c(1:152)],  col = 'navy',  lwd = 3,  type = 'l', ylab = 'Cumulative Cases', xlab = "Time (Days)",  xlim = as.Date(c("2020-01-31", "2020-06-30")), ylim = c(0,150000), main = "Number of Cumulative Cases \n with Interventions") 
#current situation
lines(italy.cases[c(10:77), 2] ~ axis_italy_p[c(1:68)], col = 'chartreuse4', lwd = 3) 
#with intervetnion
 lines(cumsum(X_italy_p[c(1:152), 5]) ~ axis_italy_p[c(1:152)],  col = 'red', lwd = 3) 
legend("topright", legend = c("Without Interventions", "Current Situation", "With Interventions"), fill = c("navy", "chartreuse4", "red"), col = c("navy", "chartreuse4", "red"), cex = 0.75, box.lty = 0)

```


####Model Application intervention a week earlier
```{r  results='hide', fig.show='hide'}

##############################Intervention###########################################

#assumption: random mixing, person to person transmission, no difference in individual immunity, no deaths
# Parameter values
N_italy <- 60480000
I_italy <- 2 
E_italy <- 20 * I_italy
R_italy <- 0 
S_italy <- N_italy-E_italy-I_italy-R_italy
alpha <- 1/5.6 #incubation period
gamma <- 1/5 #recovery period
duration <-5 #days 
# R0 and beta for different phases based on the impact of intervention to minimize transmission #using infromation from more recent papers. Since R0 = ce/D, these interventions are targeting the ce (contact rate) to reduce transmision. 


R0_1 <- 1.5 #low end (the CI from JHU and authors)
R0_2 <- 2.6 #what was calculated as gthe potential R
R0_3 <- 3.5 #high end (the CI from JHU and authors)
R0_4 <- 0.5 #updated: reduces contact by up to 85% because of lockdown
beta_1 <- R0_1/(N_italy * duration)
beta_2 <- R0_2/(N_italy * duration)
beta_3 <- R0_3/(N_italy * duration)
beta_4 <- R0_4/(N_italy * duration)

#time specification
dt <- 1
start <- 0
end_italy_p <- 151 #days
time_italy_p <- seq(start, end_italy_p, by = dt)
niter_italy_p <- (end_italy_p - start) / dt + 1
s_date_italy_p <- mdy("1-31-2020")
e_date_italy_p <- mdy("6-30-2020")
axis_italy_p <- seq(from = s_date_italy_p, to = e_date_italy_p, by = "day")


X_italy_p_e<- array(data = 0, dim = c(niter_italy_p, 5)) #created an array to catch the results
colnames(X_italy_p_e) <- c("S", "E", "I","R", "New Cases")

X_italy_p_e[1, 1] <- S_italy #initial susuceptible population 
X_italy_p_e[1, 2] <- E_italy #initial latent population 
X_italy_p_e[1, 3] <- I_italy #initial infectious population 
X_italy_p_e[1, 4] <- R_italy #initial recovered population 
X_italy_p_e[1, 5] <- I_italy #new cases

for (i in 2:niter_italy_p) { 
  if(i < 18) {
    #phase 1 transmission with an R0 of 3.1 (January 31 - February 23rd)
    X_italy_p_e[i,1] <- X_italy_p_e[i-1, 1] + dt * (-beta_3 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3] )  #S
    X_italy_p_e[i,2]<- X_italy_p_e[i-1,2] + dt * ((beta_3 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3]) - alpha * X_italy_p_e[i-1,2]) #E
    X_italy_p_e[i,3] <- X_italy_p_e[i-1,3] + dt * (alpha * X_italy_p_e[i-1,2] - gamma * X_italy_p_e[i-1,3])  #I 
    X_italy_p_e[i,4] <- X_italy_p_e[i-1,4] + dt * (gamma * X_italy_p_e[i-1,3]) #R
    X_italy_p_e[i,5] <- dt * (alpha * X_italy_p_e[i-1,2])
  }
  else {
    if ( i >= 18 & i < 34) { 
      #phase 2 transmission with an R0 of 1.9 (February 24 -  March 8) blocking the north
      X_italy_p_e[i,1] <- X_italy_p_e[i-1, 1] + dt * (-beta_2 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3] )  #S
      X_italy_p_e[i,2]<- X_italy_p_e[i-1,2] + dt * ((beta_2 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3]) - alpha * X_italy_p_e[i-1,2]) #E
      X_italy_p_e[i,3] <- X_italy_p_e[i-1,3] + dt * (alpha * X_italy_p_e[i-1,2] - gamma * X_italy_p_e[i-1,3])  #I 
      X_italy_p_e[i,4] <- X_italy_p_e[i-1,4] + dt * (gamma * X_italy_p_e[i-1,3]) #R
      X_italy_p_e[i,5] <- dt * (alpha * X_italy_p_e[i-1,2])
    }
    else{
      if( i >= 34 & i < 48){
        #phase 3 transmission with an R0 of 0.65 (March 8 - March 10) North total Lockdown
        X_italy_p_e[i,1] <- X_italy_p_e[i-1, 1] + dt * (-beta_1 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3] )  #S
        X_italy_p_e[i,2]<- X_italy_p_e[i-1,2] + dt * ((beta_1 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3]) - alpha * X_italy_p_e[i-1,2]) #E
        X_italy_p_e[i,3] <- X_italy_p_e[i-1,3] + dt * (alpha * X_italy_p_e[i-1,2] - gamma * X_italy_p_e[i-1,3])  #I 
        X_italy_p_e[i,4] <- X_italy_p_e[i-1,4] + dt * (gamma * X_italy_p_e[i-1,3]) #R
        X_italy_p_e[i,5] <- dt * (alpha * X_italy_p_e[i-1,2])  
      }
      else{
        X_italy_p_e[i,1] <- X_italy_p_e[i-1, 1] + dt * (-beta_4 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3])  #S
        X_italy_p_e[i,2]<- X_italy_p_e[i-1,2] + dt * ((beta_4 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3]) - alpha * X_italy_p_e[i-1,2]) #E
        X_italy_p_e[i,3] <- X_italy_p_e[i-1,3] + dt * (alpha * X_italy_p_e[i-1,2] - gamma * X_italy_p_e[i-1,3])  #I 
        X_italy_p_e[i,4] <- X_italy_p_e[i-1,4] + dt * (gamma * X_italy_p_e[i-1,3]) #R
        X_italy_p_e[i,5] <- dt * (alpha * X_italy_p_e[i-1,2])  
      }
    }
  }
}

X_italy_p_e<- as.data.frame(X_italy_p_e)
X_italy_p_e$Date <- axis_italy_p
max(X_italy_p_e$I)



###########plots#################

#plot: number of new cases main = "Impact of Intervention on the Number of Infectious People Each Day",
plot(X_italy_p_e[c(1:18), 3] ~ axis_italy_p[c(1:18)] , type = 'l', col = 'steelblue4',  lwd = 3,  ylab = "Number of People Infectious", xlab = "Time (Days)",  xlim = as.Date(c("2020-01-31", "2020-06-30")), ylim = c(0,10000), main = "Number of People Infectious \n with Interventions")
lines(X_italy_p_e[c(18:34), 3] ~ axis_italy_p[c(18:34)] , col = 'violetred4', lwd = 3)
lines(X_italy_p_e[c(34:48), 3] ~ axis_italy_p[c(34:48)], col = 'gold', lwd = 3)
lines(X_italy_p_e[c(48:152), 3] ~ axis_italy_p[c(48:152)], col = 'red', lwd = 3)
lines(X_italy_p[c(24:41), 3] ~ axis_italy_p[c(24:41)] , col = 'violetred4', lwd = 3)
lines(X_italy_p[c(41:55), 3] ~ axis_italy_p[c(41:55)], col = 'gold', lwd = 3)
lines(X_italy_p[c(55:152), 3] ~ axis_italy_p[c(55:152)], col = 'red', lwd = 3)
lines(X_italy_p_n[c(1:152), 3] ~ axis_italy_p[c(1:152)] , col = 'navy', lwd = 3) #if we are assuming that transmission continued at a high rate
lines(X_italy_p[c(1:24), 3] ~ axis_italy_p[c(1:24)] ,  col = 'steelblue4',  lwd = 3)

legend("topright", legend = c("R=3.5", "R=2.6", "R=1.5", "R=0.5", "No Intervention"), fill = c("steelblue4", "violetred4", "gold", "red", "navy"), col = c("steelblue4", "violetred4", "gold", "red", "navy"), cex = 0.75, box.lty = 0)



#cumulative cases comparison
#intervention: main = "Comparing the Model Interventions with Non-intervention",
plot(cumsum(X_italy_p_e[c(1:152), 5]) ~ axis_italy_p[c(1:152)] , type = 'l', col = 'pink',  lwd = 3,  ylab = 'Cumulative Cases', xlab = "Time (Days)",  xlim = as.Date(c("2020-01-31", "2020-06-30")), ylim = c(0,150000), main = "Number of Cumulative Cases \n with Interventions") #with intervetnion
 lines(cumsum(X_italy_p_n[c(1:152), 5]) ~ axis_italy_p[c(1:152)] , col = 'navy', lwd = 3) 
lines(italy.cases[c(10:77), 2] ~ axis_italy_p[c(1:68)], col = 'chartreuse4', lwd = 3) 
lines(cumsum(X_italy_p[c(1:152), 5]) ~ axis_italy_p[c(1:152)] ,  col = 'red',  lwd = 3)
legend("topright", legend = c("Without Interventions", "Current Situation", "With Interventions", "Intervention A \n week earlier"), fill = c("navy", "chartreuse4", "red","pink" ), col = c("navy", "chartreuse4", "red","pink" ), cex = 0.75, box.lty = 0)

```

####Comparing Values if intervention is earlier or at current date
```{r  results='hide', fig.show='hide'}
max.nointervention <- max(X_italy_p_n$I)

max.phase.infect <- max(X_italy_p$I)
cum.phase.infect <- max(cumsum(X_italy_p$`New Cases`))

max.phase.infect.early <- max(X_italy_p_e$I)
cum.phase.infect.early <- max(cumsum(X_italy_p_e$`New Cases`))
```

