---
title: "Italy with Deaths"
author: "Afia Amoako"
date: '2020-03-31'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#required package
library(httr)
library(tidyverse)
library(lubridate)
library(knitr)
```


##Introduction

Analysis for the impact of intervention on COVID transmission is adapted from Wang, H., Wang, Z., Dong, Y., Chang, R., Xu, C., Yu, X., Zhang, S., Tsamlag, L., Shang, M., Huang, J., Wang, Y., Xu, G., Shen, T., Zhang, X., & Cai, Y. (2020). Phase-adjusted estimation of the number of Coronavirus Disease 2019 cases in Wuhan, China. Cell Discovery, 6(1), 1–8. https://doi.org/10.1038/s41421-020-0148-0


###Importing Data 
```{r}
COVID <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
head(COVID, 10)
summary(COVID)

COVID_deaths <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")
head(COVID_deaths)
summary(COVID_deaths)

#things to do, downloads deaths, add death to all models, plot deaths 
```

###Cleaning Italy Table 
```{r}
#analysis up to 28th march 

#CASES
#cleaning removing province, longititude and latitude 
COVID.clean <- COVID[ ,-c(1,3,4)]
#subset Italy cumulative cases
COVID.italy <- COVID.clean %>% filter(Country.Region == "Italy")
#creating cumulative cases table for  Italy 
start_date <- mdy("1-22-2020")
end_date <- mdy("3-31-2020")
intervals <- seq(from = start_date, to=end_date, by = "day")
cases <- as.numeric(COVID.italy[1, -1])
cum.cases.italy <- as.data.frame(cbind(Date = intervals, Cases = cases))
cum.cases.italy$Date <- as.Date(cum.cases.italy$Date, origin ="1970-01-01")



#DEATHS
COVID.clean.deaths <- COVID_deaths[ ,-c(1,3,4)]
#subset Italy cumulative deaths
COVID.italy.deaths <- COVID.clean.deaths %>% filter(Country.Region == "Italy")
#creating cumulative cases table for  Italy 
deaths <- as.numeric(COVID.italy.deaths[1, -1])
cum.deaths.italy <- as.data.frame(cbind(Date = intervals, Deaths = deaths))
cum.deaths.italy$Date <- as.Date(cum.deaths.italy$Date, origin ="1970-01-01")

#plot.cumulative cases and Deaths 
plot(cum.cases.italy$Date, cum.cases.italy$Cases, type = "l", lwd = "3", col = "navy", main = "Cumulative Cases in Italy", xlab = "Date", ylab = "Cases", ylim = c(0,100000))
lines(cum.deaths.italy$Date, cum.deaths.italy$Deaths, lwd = "3",   col = "chartreuse4")
legend("topleft", legend = c( "Cases","Deaths"), fill = c("navy", "chartreuse4"), col = c("navy", "chartreuse4"), cex = 0.5)
```

#Italy
##Model Application: Estimating the potential number of infectious people based on different R0 values
```{r}
#assumption: random mixing, person to person transmission, no difference in individual immunity, no deaths

# Parameter values
#initial variables per year
N_italy <- 60480000 #population of Italy
I_italy <- 2 #ikmported cases from Italy
E_italy <- 20 * I_italy #number potentially exposed
R_italy <- 0 
S_italy <- N_italy-E_italy-I_italy-R_italy
br <- (7.6/1000)/365
mu_pop <- (10.7/1000)/365
mu_c <- 5/100
alpha <- 1/5.6 #incubation period
gamma <- 1/5 #recovery period
duration <- 5 #days 

# R0 and beta for different phases based on the impact of intervention to minimize transmission
R0_1 <- 1.5 
R0_2 <- 2.6
R0_3 <- 3.5 #without any intervention
R0_4 <- 0.9
beta_1 <- R0_1/(N_italy*duration)
beta_2 <- R0_2/(N_italy*duration)
beta_3 <- R0_3/(N_italy*duration)
beta_4 <- R0_4/(N_italy*duration)

#time specification
dt <- 1
start <- 0
end_italy <- 40
time_italy <- seq(start, end_italy, by = dt)
niter_italy <- (end_italy - start) / dt + 1
s_date_italy <- mdy("1-31-2020") #first cases reported in Rome
e_date_italy <- mdy("3-11-2020") #last up to date results
axis_italy <- seq(from = s_date_italy, to = e_date_italy, by = "day")


#create table
l <- 1 #index for beta_1: R0: 1.9
m <- 2 #index for beta_2: R0: 2.6
h <- 3 #index for beta_2: R0: 3.5

X_italy_d <- array(data = 0, dim = c(niter_italy, 6, 3)) #created an array to catch the results
colnames(X_italy_d) <- c("S", "E", "I","R", "New Case", "Total Deaths")
#low
X_italy_d[1, 1, l] <- S_italy #initial susuceptible population R0: 1.9
X_italy_d[1, 2, l] <- E_italy #initial latent population R0: 1.9
X_italy_d[1, 3, l] <- I_italy #initial infectious population R0: 1.9
X_italy_d[1, 4, l] <- R_italy #initial recovered population R0: 1.9
X_italy_d[1, 5, l] <- I_italy #initial infectious population R0: 1.9: using this row to count the number of new cases for cumulative incidence calculation.
X_italy_d[1, 6, l] <- 0 #initial dead

#medium
X_italy_d[1, 1, m] <- S_italy #initial susuceptible population R0: 2.6
X_italy_d[1, 2, m] <- E_italy #initial latent population R0: 2.6
X_italy_d[1, 3, m] <- I_italy #initial infectious population R0: 2.6
X_italy_d[1, 4, m] <- R_italy #initial recovered population R0: 2.6
X_italy_d[1, 5, m] <- I_italy #initial infectious population R0: 2.6: using this row to count the number of new cases for cumulative incidence calculation.
X_italy_d[1, 6, m] <- 0 #initial dead


#high 
X_italy_d[1, 1, h] <- S_italy #initial susuceptible population R0: 3.5
X_italy_d[1, 2, h] <- E_italy #initial latent population R0: 3.5
X_italy_d[1, 3, h] <- I_italy #initial infectious population R0: 3.5
X_italy_d[1, 4, h] <- R_italy #initial recovered population R0: 3.5
X_italy_d[1, 5, h] <- I_italy #initial infectious population R0: 3.5: using this row to count the number of new cases for cumulative incidence calculation.
X_italy_d[1, 6, h] <- 0 #initial dead


for (i in 2:niter_italy) { 
  #low
  X_italy_d[i,1, l] <- X_italy_d[i-1, 1, l] + dt * (-beta_1 * X_italy_d[i-1,1,l] * X_italy_d[i-1,3,l] - mu_pop * X_italy_d[i-1,1,l] + br * sum(X_italy_d[i-1,c(1:4) ,l]) )  #S
  X_italy_d[i,2, l]<- X_italy_d[i-1,2, l] + dt * (beta_1 * X_italy_d[i-1,1,l] *  X_italy_d[i-1,3,l] - (alpha + mu_pop) * X_italy_d[i-1,2,l]) #E
  X_italy_d[i,3, l] <- X_italy_d[i-1,3, l] + dt * (alpha * X_italy_d[i-1,2, l] - (gamma + mu_pop) * X_italy_d[i-1,3, l] - mu_c * X_italy_d[i-1,3, l])  #I 
  X_italy_d[i,4, l] <- X_italy_d[i-1,4, l] + dt * (gamma * X_italy_d[i-1,3, l] - mu_pop * X_italy_d[i-1,4, l]) #R
  X_italy_d[i,5, l] <- dt * (alpha * X_italy_d[i-1,2, l]) #new cases added 
  X_italy_d[i,6, l] <- dt * ((mu_c * X_italy_d[i-1,3, l]) + (mu_pop * (X_italy_d[i-1,1,l] + X_italy_d[i-1,2,l] + X_italy_d[i-1,3,l] + X_italy_d[i-1,4, l])))
  
  #medium
  X_italy_d[i,1, m] <- X_italy_d[i-1, 1, m] + dt * (-beta_2 * X_italy_d[i-1,1,m] * X_italy_d[i-1,3,m] - mu_pop * X_italy_d[i-1,1,m] + br * sum(X_italy_d[i-1, ,m]))  #S
  X_italy_d[i,2, m]<- X_italy_d[i-1,2, m] + dt * (beta_2 * X_italy_d[i-1,1,m] *  X_italy_d[i-1,3,m] - (alpha + mu_pop) * X_italy_d[i-1,2,m]) #E
  X_italy_d[i,3, m] <- X_italy_d[i-1,3, m] + dt * (alpha * X_italy_d[i-1,2, m] - (gamma + mu_pop) * X_italy_d[i-1,3, m] - mu_c * X_italy_d[i-1,3, m])  #I 
  X_italy_d[i,4, m] <- X_italy_d[i-1,4, m] + dt * (gamma * X_italy_d[i-1,3, m] - mu_pop * X_italy_d[i-1,4, m]) #R
  X_italy_d[i,5, m] <- dt * (alpha * X_italy_d[i-1,2, m]) #new cases added 
   X_italy_d[i,6, m] <- dt * ((mu_c * X_italy_d[i-1,3, m]) + (mu_pop * (X_italy_d[i-1,1,m] + X_italy_d[i-1,2,m] + X_italy_d[i-1,3,m] + X_italy_d[i-1,4, m])))
   
  #high
  X_italy_d[i,1, h] <- X_italy_d[i-1, 1, h] + dt * (-beta_3 * X_italy_d[i-1,1,h] * X_italy_d[i-1,3,h] - mu_pop * X_italy_d[i-1,1,h] + br * sum(X_italy_d[i-1, ,m]))  #S
  X_italy_d[i,2, h]<- X_italy_d[i-1,2, h] + dt * (beta_3 * X_italy_d[i-1,1,h] *  X_italy_d[i-1,3,h] - (alpha + mu_pop) * X_italy_d[i-1,2,h]) #E
  X_italy_d[i,3, h] <- X_italy_d[i-1,3, h] + dt * (alpha * X_italy_d[i-1,2, h] - (gamma + mu_pop) * X_italy_d[i-1,3, h] - mu_c * X_italy_d[i-1,3, h])  #I 
  X_italy_d[i,4, h] <- X_italy_d[i-1,4, h] + dt * (gamma * X_italy_d[i-1,3, h] - mu_pop * X_italy_d[i-1,4, h]) #R
  X_italy_d[i,5, h] <- dt * (alpha * X_italy_d[i-1,2, h]) #new cases added 
    X_italy_d[i,6, h] <- dt * ((mu_c * X_italy_d[i-1,3, h]) + (mu_pop * (X_italy_d[i-1,1,h] + X_italy_d[i-1,2,h] + X_italy_d[i-1,3,h] + X_italy_d[i-1,4, h])))
  
}

low_beta_italy_d <- as.data.frame(X_italy_d[ , ,l])
medium_beta_italy_d <- as.data.frame(X_italy_d[ , ,m])
high_beta_italy_d <- as.data.frame(X_italy_d[ , ,h])
high_beta_italy_d$Date <- axis_italy
medium_beta_italy_d$Date <- axis_italy
low_beta_italy_d$Date <- axis_italy


#"Number of People Infectious in  Italy with varying R0 values",
plot(high_beta_italy_d$I ~ axis_italy, type = 'l',  xlab = "Date", ylab = "Number of People Infectious", col = 'navy', lwd = 3, ylim = c(0,5000))
 lines(medium_beta_italy_d$I ~ axis_italy, col = 'red', lwd = 3)
 lines(low_beta_italy_d$I ~ axis_italy, col = 'pink', lwd = 3)
 legend("topleft", legend = c( "Model R: 3.5","Model R:2.6", "Model R: 1.5"), fill = c("navy", "red", "pink"), col = c("navy", "red", "pink"), cex = 0.5)

#Number of People Dying 
plot(high_beta_italy_d$`Total Deaths` ~ axis_italy, type = 'l',  xlab = "Date", ylab = "Number of People Dead", col = 'navy', lwd = 3)
 lines(medium_beta_italy_d$`Total Deaths` ~ axis_italy, col = 'red', lwd = 3)
 lines(low_beta_italy_d$`Total Deaths` ~ axis_italy, col = 'pink', lwd = 3)
 legend("topleft", legend = c( "Model R: 3.5","Model R:2.6", "Model R: 1.5"), fill = c("navy", "red", "pink"), col = c("navy", "red", "pink"), cex = 0.5)
 
 
  
#"Comparing Modeled Cumulative Cases in Italy with varying R0 values to current cases",

plot(cum.cases.italy[c(10:50), 1], cum.cases.italy[c(10:50), 2], type = 'l', xlab = "Date", ylab = "Cumulative Cases", col = "chartreuse4", lwd = 3)
 lines(cumsum(low_beta_italy_d$`New Case`) ~ axis_italy, col = "pink", lwd = 3)
 lines(cumsum(medium_beta_italy_d$`New Case`) ~ axis_italy, col = 'red', lwd = 3)
 lines(cumsum(high_beta_italy_d$`New Case`) ~ axis_italy, col = "navy", lwd = 3)
 legend("topleft", legend = c("Current Cases \n in Italy", "Model R: 3.5","Model R:2.6", "Model R:1.5"), fill = c("chartreuse4", "navy", "red", "pink"), col = c("chartreuse4", "navy", "red", "pink"), cex = 0.5)
  #caption to see how my model can relate to what is currently seen in italy, I used the model to calculate the number of new infectious people occuring based on the number of people moving into the I column and compared this to the current information on cumulative cases. This will allow me to compare with current cases. 
 
 #"Comparing Modeled Cumulative Cases in Italy with varying R0 values to current cases",
plot(cum.deaths.italy[c(10:50), 1], cum.deaths.italy[c(10:50), 2], type = 'l', xlab = "Date", ylab = "Cumulative Deaths", col = "chartreuse4", lwd = 3)
 lines(cumsum(low_beta_italy_d$`Total Deaths`) ~ axis_italy, col = "pink", lwd = 3)
 lines(cumsum(medium_beta_italy_d$`Total Deaths`) ~ axis_italy, col = 'red', lwd = 3)
 lines(cumsum(high_beta_italy_d$`Total Deaths`) ~ axis_italy, col = "navy", lwd = 3)
 legend("topleft", legend = c("Current Deaths \n in Italy", "Model R: 3.5","Model R:2.6", "Model R:1.5"), fill = c("chartreuse4", "navy", "red", "pink"), col = c("chartreuse4", "navy", "red", "pink"), cex = 0.5)
 

 


 
#after graphing, it seem that 3.5 is actually a better growth rate because it is similar to the italy number. Assuming that italy is only counting those who are presenting for testing, tends to be those with some symptoms and this is not including those who are incubating, so could be under estimation. Thus used the I to represent the symptomatic and calculated cumulative cases by the number of cases moving from E to I each time step and then taking the  cumulative sum. 3.5 is a good R0 to start with. 
```
