---
title: "Italy Analysis"
author: "Afia Amoako"
date: '2020-03-29'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#required package
library(httr)
library(tidyverse)
library(lubridate)
library(knitr)
```

##Introduction

Analysis for the impact of intervention on COVID transmission is adapted from Wang, H., Wang, Z., Dong, Y., Chang, R., Xu, C., Yu, X., Zhang, S., Tsamlag, L., Shang, M., Huang, J., Wang, Y., Xu, G., Shen, T., Zhang, X., & Cai, Y. (2020). Phase-adjusted estimation of the number of Coronavirus Disease 2019 cases in Wuhan, China. Cell Discovery, 6(1), 1–8. https://doi.org/10.1038/s41421-020-0148-0


###Importing Data 
```{r}
COVID <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
head(COVID, 10)
summary(COVID)

#things to do, downloads deaths, add death to all models, plot deaths 
```

###Cleaning Italy Table 
```{r}
#analysis up to 28th march 

#CASES
#cleaning removing province, longititude and latitude 
COVID.clean <- COVID[ ,-c(1,3,4)]

#subset Italy cumulative cases
COVID.italy <- COVID.clean %>% filter(Country.Region == "Italy")
#creating cumulative cases table for  Italy 
start_date <- mdy("1-22-2020")
end_date <- mdy("3-31-2020")
intervals <- seq(from = start_date, to=end_date, by = "day")
cases <- as.numeric(COVID.italy[1, -1])
cum.cases.italy <- as.data.frame(cbind(Date = intervals, Cases = cases))
cum.cases.italy$Date <- as.Date(cum.cases.italy$Date, origin ="1970-01-01")

#plot.cummulative.cases
plot(cum.cases.italy$Date, type = "l", lwd = "3", col = "navy", cum.cases.italy$Cases, main = "Cumulative Cases in Italy", xlab = "Date", ylab = "Cases")
```

#Italy
##Model Application: Estimating the potential number of infectious people based on different R0 values
```{r}
#assumption: random mixing, person to person transmission, no difference in individual immunity, no deaths

# Parameter values
#initial variables per year
N_italy <- 60480000 #population of Italy
I_italy <- 2 #ikmported cases from Italy
E_italy <- 20 * I_italy #number potentially exposed
R_italy <- 0 
S_italy <- N_italy-E_italy-I_italy-R_italy
alpha <- 1/5.6 #incubation period
gamma <- 1/5 #recovery period
duration <- 5 #days 

# R0 and beta for different phases based on the impact of intervention to minimize transmission
R0_1 <- 1.5 
R0_2 <- 2.6
R0_3 <- 3.5 #without any intervention
R0_4 <- 0.9
beta_1 <- R0_1/(N_italy*duration)
beta_2 <- R0_2/(N_italy*duration)
beta_3 <- R0_3/(N_italy*duration)
beta_4 <- R0_4/(N_italy*duration)

#time specification
dt <- 1
start <- 0
end_italy <- 40
time_italy <- seq(start, end_italy, by = dt)
niter_italy <- (end_italy - start) / dt + 1
s_date_italy <- mdy("1-31-2020") #first cases reported in Rome
e_date_italy <- mdy("3-11-2020") #last up to date results
axis_italy <- seq(from = s_date_italy, to = e_date_italy, by = "day")


#create table
l <- 1 #index for beta_1: R0: 1.9
m <- 2 #index for beta_2: R0: 2.6
h <- 3 #index for beta_2: R0: 3.5

X_italy <- array(data = 0, dim = c(niter_italy, 5, 3)) #created an array to catch the results
colnames(X_italy) <- c("S", "E", "I","R", "New Case")
#low
X_italy[1, 1, l] <- S_italy #initial susuceptible population R0: 1.9
X_italy[1, 2, l] <- E_italy #initial latent population R0: 1.9
X_italy[1, 3, l] <- I_italy #initial infectious population R0: 1.9
X_italy[1, 4, l] <- R_italy #initial recovered population R0: 1.9
X_italy[1, 5, l] <- I_italy #initial infectious population R0: 1.9: using this row to count the number of new cases for cumulative incidence calculation.

#medium
X_italy[1, 1, m] <- S_italy #initial susuceptible population R0: 2.6
X_italy[1, 2, m] <- E_italy #initial latent population R0: 2.6
X_italy[1, 3, m] <- I_italy #initial infectious population R0: 2.6
X_italy[1, 4, m] <- R_italy #initial recovered population R0: 2.6
X_italy[1, 5, m] <- I_italy #initial infectious population R0: 2.6: using this row to count the number of new cases for cumulative incidence calculation.


#high 
X_italy[1, 1, h] <- S_italy #initial susuceptible population R0: 3.5
X_italy[1, 2, h] <- E_italy #initial latent population R0: 3.5
X_italy[1, 3, h] <- I_italy #initial infectious population R0: 3.5
X_italy[1, 4, h] <- R_italy #initial recovered population R0: 3.5
X_italy[1, 5, h] <- I_italy #initial infectious population R0: 3.5: using this row to count the number of new cases for cumulative incidence calculation.



for (i in 2:niter_italy) { 
  #low
  X_italy[i,1, l] <- X_italy[i-1, 1, l] + dt * (-beta_1 * X_italy[i-1,1,l] * (X_italy[i-1,3,l]) )  #S
  X_italy[i,2, l]<- X_italy[i-1,2, l] + dt * (beta_1 * X_italy[i-1,1,l] *  (X_italy[i-1,3,l]) - alpha * X_italy[i-1,2,l]) #E
  X_italy[i,3, l] <- X_italy[i-1,3, l] + dt * (alpha * X_italy[i-1,2, l] - gamma * X_italy[i-1,3, l])  #I 
  X_italy[i,4, l] <- X_italy[i-1,4, l] + dt * (gamma * X_italy[i-1,3, l]) #R
  X_italy[i,5, l] <- dt * (alpha * X_italy[i-1,2, l]) #new cases added 
  
  #medium
  X_italy[i,1, m] <- X_italy[i-1, 1, m] + dt * (-beta_2 * X_italy[i-1,1,m] * (X_italy[i-1,3,m])  ) #S
  X_italy[i,2, m]<- X_italy[i-1,2, m] + dt * (beta_2 * X_italy[i-1,1,m] * (X_italy[i-1,3,m]) - alpha * X_italy[i-1,2,m]) #E
  X_italy[i,3, m] <- X_italy[i-1,3, m] + dt * (alpha * X_italy[i-1,2, m] - gamma * X_italy[i-1,3, m])  #I 
  X_italy[i,4, m] <- X_italy[i-1,4, m] + dt * (gamma * X_italy[i-1,3, m]) #R    
  X_italy[i,5, m] <- dt * (alpha * X_italy[i-1,2, m]) #new cases added
   
  #high
  X_italy[i,1, h] <- X_italy[i-1, 1, h] + dt * (-beta_3 * X_italy[i-1,1,h] * (X_italy[i-1,3,h]) ) #S
  X_italy[i,2, h]<- X_italy[i-1,2, h] + dt * (beta_3 * X_italy[i-1,1,h] * (X_italy[i-1,3,h]) - alpha * X_italy[i-1,2,h]) #E
  X_italy[i,3, h] <- X_italy[i-1,3, h] + dt * (alpha * X_italy[i-1,2, h] - gamma * X_italy[i-1,3, h])  #I 
  X_italy[i,4, h] <- X_italy[i-1,4, h] + dt * (gamma * X_italy[i-1,3, h]) #R
  X_italy[i,5, h] <- dt * (alpha * X_italy[i-1,2, h]) #new cases added
  
}

low_beta_italy <- as.data.frame(X_italy[ , ,l])
medium_beta_italy <- as.data.frame(X_italy[ , ,m])
high_beta_italy <- as.data.frame(X_italy[ , ,h])
high_beta_italy$Date <- axis_italy
high_beta_italy$CumSum <- cumsum(high_beta_italy$`New Case`)
medium_beta_italy$Date <- axis_italy
low_beta_italy$Date <- axis_italy


#"Number of People Infectious in  Italy with varying R0 values",
par(mfrow = c(1, 2))
plot(high_beta_italy$I ~ axis_italy, type = 'l',  xlab = "Date", ylab = "Number of People Infectious", col = 'navy', lwd = 3)
 lines(medium_beta_italy$I ~ axis_italy, col = 'red', lwd = 3)
 lines(low_beta_italy$I ~ axis_italy, col = 'pink', lwd = 3)
 legend("topleft", legend = c( "Model R: 3.5","Model R:2.6", "Model R: 1.5"), fill = c("navy", "red", "pink"), col = c("navy", "red", "pink"), cex = 0.5)

 
#"Comparing Modeled Cumulative Cases in Italy with varying R0 values to current cases",
plot(cum.cases.italy[c(10:50), 1], cum.cases.italy[c(10:50), 2], type = 'l', xlab = "Date", ylab = "Cumulative Cases", col = "chartreuse4", lwd = 3)
 lines(cumsum(low_beta_italy$`New Case`) ~ axis_italy, col = "pink", lwd = 3)
 lines(cumsum(medium_beta_italy$`New Case`) ~ axis_italy, col = 'red', lwd = 3)
 lines(cumsum(high_beta_italy$`New Case`) ~ axis_italy, col = "navy", lwd = 3)
 legend("topleft", legend = c("Current Cases \n in Italy", "Model R: 3.5","Model R:2.6", "Model R:1.5"), fill = c("chartreuse4", "navy", "red", "pink"), col = c("chartreuse4", "navy", "red", "pink"), cex = 0.5)
  #caption to see how my model can relate to what is currently seen in italy, I used the model to calculate the number of new infectious people occuring based on the number of people moving into the I column and compared this to the current information on cumulative cases. This will allow me to compare with current cases. 
 
 par(mfrow = c(1, 1))

 
#after graphing, it seem that 3.5 is actually a better growth rate because it is similar to the italy number. Assuming that italy is only counting those who are presenting for testing, tends to be those with some symptoms and this is not including those who are incubating, so could be under estimation. Thus used the I to represent the symptomatic and calculated cumulative cases by the number of cases moving from E to I each time step and then taking the  cumulative sum. 3.5 is a good R0 to start with. 
```

##Model Application intervention!

I applied interventions put in place to slow transmission looked at how that will change incidences. I compared my values to the cumulative cases, looking at the current situation, my model approximation of the current situation and if the intervention actually creates an impact

```{r}

##############################Intervention###########################################

#assumption: random mixing, person to person transmission, no difference in individual immunity, no deaths
# Parameter values
N_italy <- 60480000
I_italy <- 2 
E_italy <- 20 * I_italy
R_italy <- 0 
S_italy <- N_italy-E_italy-I_italy-R_italy
alpha <- 1/5.6 #incubation period
gamma <- 1/5 #recovery period
duration <-5 #days 
# R0 and beta for different phases based on the impact of intervention to minimize transmission #using infromation from more recent papers. Since R0 = ce/D, these interventions are targeting the ce (contact rate) to reduce transmision. 


R0_1 <- 1.5 #low end (the CI from JHU and authors)
R0_2 <- 2.6 #what was calculated as gthe potential R
R0_3 <- 3.5 #high end (the CI from JHU and authors)
R0_4 <- 0.5 #updated: reduces contact by up to 85% because of lockdown
beta_1 <- R0_1/(N_italy * duration)
beta_2 <- R0_2/(N_italy * duration)
beta_3 <- R0_3/(N_italy * duration)
beta_4 <- R0_4/(N_italy * duration)

#time specification
dt <- 1
start <- 0
end_italy_p <- 90 #days
time_italy_p <- seq(start, end_italy_p, by = dt)
niter_italy_p <- (end_italy_p - start) / dt + 1
s_date_italy_p <- mdy("1-31-2020")
e_date_italy_p <- mdy("4-30-2020")
axis_italy_p <- seq(from = s_date_italy_p, to = e_date_italy_p, by = "day")


X_italy_p <- array(data = 0, dim = c(niter_italy_p, 5)) #created an array to catch the results
colnames(X_italy_p) <- c("S", "E", "I","R", "New Cases")

X_italy_p[1, 1] <- S_italy #initial susuceptible population 
X_italy_p[1, 2] <- E_italy #initial latent population 
X_italy_p[1, 3] <- I_italy #initial infectious population 
X_italy_p[1, 4] <- R_italy #initial recovered population 
X_italy_p[1, 5] <- I_italy #new cases

for (i in 2:niter_italy_p) { 
  if(i < 25) {
  #phase 1 transmission with an R0 of 3.1 (January 31 - February 23rd)
  X_italy_p[i,1] <- X_italy_p[i-1, 1] + dt * (-beta_3 * X_italy_p[i-1,1] * X_italy_p[i-1,3] )  #S
  X_italy_p[i,2]<- X_italy_p[i-1,2] + dt * ((beta_3 * X_italy_p[i-1,1] * X_italy_p[i-1,3]) - alpha * X_italy_p[i-1,2]) #E
  X_italy_p[i,3] <- X_italy_p[i-1,3] + dt * (alpha * X_italy_p[i-1,2] - gamma * X_italy_p[i-1,3])  #I 
  X_italy_p[i,4] <- X_italy_p[i-1,4] + dt * (gamma * X_italy_p[i-1,3]) #R
  X_italy_p[i,5] <- dt * (alpha * X_italy_p[i-1,2])
  }
  else {
    if ( i >= 25 & i < 41) { 
  #phase 2 transmission with an R0 of 1.9 (February 24 -  March 8) blocking the north
  X_italy_p[i,1] <- X_italy_p[i-1, 1] + dt * (-beta_2 * X_italy_p[i-1,1] * X_italy_p[i-1,3] )  #S
  X_italy_p[i,2]<- X_italy_p[i-1,2] + dt * ((beta_2 * X_italy_p[i-1,1] * X_italy_p[i-1,3]) - alpha * X_italy_p[i-1,2]) #E
  X_italy_p[i,3] <- X_italy_p[i-1,3] + dt * (alpha * X_italy_p[i-1,2] - gamma * X_italy_p[i-1,3])  #I 
  X_italy_p[i,4] <- X_italy_p[i-1,4] + dt * (gamma * X_italy_p[i-1,3]) #R
  X_italy_p[i,5] <- dt * (alpha * X_italy_p[i-1,2])
    }
    else{
      if( i >= 41 & i < 55){
        #phase 3 transmission with an R0 of 0.65 (March 8 - March 10) North total Lockdown
        X_italy_p[i,1] <- X_italy_p[i-1, 1] + dt * (-beta_1 * X_italy_p[i-1,1] * X_italy_p[i-1,3] )  #S
        X_italy_p[i,2]<- X_italy_p[i-1,2] + dt * ((beta_1 * X_italy_p[i-1,1] * X_italy_p[i-1,3]) - alpha * X_italy_p[i-1,2]) #E
         X_italy_p[i,3] <- X_italy_p[i-1,3] + dt * (alpha * X_italy_p[i-1,2] - gamma * X_italy_p[i-1,3])  #I 
        X_italy_p[i,4] <- X_italy_p[i-1,4] + dt * (gamma * X_italy_p[i-1,3]) #R
         X_italy_p[i,5] <- dt * (alpha * X_italy_p[i-1,2])  
      }
      else{
        X_italy_p[i,1] <- X_italy_p[i-1, 1] + dt * (-beta_4 * X_italy_p[i-1,1] * X_italy_p[i-1,3])  #S
        X_italy_p[i,2]<- X_italy_p[i-1,2] + dt * ((beta_4 * X_italy_p[i-1,1] * X_italy_p[i-1,3]) - alpha * X_italy_p[i-1,2]) #E
         X_italy_p[i,3] <- X_italy_p[i-1,3] + dt * (alpha * X_italy_p[i-1,2] - gamma * X_italy_p[i-1,3])  #I 
        X_italy_p[i,4] <- X_italy_p[i-1,4] + dt * (gamma * X_italy_p[i-1,3]) #R
         X_italy_p[i,5] <- dt * (alpha * X_italy_p[i-1,2])  
      }
      }
  }
  }

X_italy_p <- as.data.frame(X_italy_p)
X_italy_p$date <- axis_italy_p

  
##############################Comparison with no iuntervention###########################################
#no intervention using high transmission rate 
X_italy_p_n <- array(data = 0, dim = c(niter_italy_p, 5)) #created an array to catch the results
colnames(X_italy_p_n) <- c("S", "E", "I","R", "New Cases")

X_italy_p_n[1, 1] <- S_italy #initial susuceptible population 
X_italy_p_n[1, 2] <- E_italy #initial latent population 
X_italy_p_n[1, 3] <- I_italy #initial infectious population 
X_italy_p_n[1, 4] <- R_italy #initial recovered population 
X_italy_p_n[1, 5] <- I_italy #new cases

for(i in 2:niter_italy_p) {
  X_italy_p_n [i,1] <- X_italy_p_n [i-1, 1] + dt * (-beta_3 * X_italy_p_n [i-1,1] * X_italy_p_n[i-1,3] )  #S
  X_italy_p_n [i,2]<- X_italy_p_n [i-1,2] + dt * ((beta_3 * X_italy_p_n [i-1,1] * X_italy_p_n[i-1,3]) - alpha * X_italy_p[i-1,2]) #E
  X_italy_p_n [i,3] <- X_italy_p_n [i-1,3] + dt * (alpha * X_italy_p_n [i-1,2] - gamma * X_italy_p_n [i-1,3])  #I 
  X_italy_p_n [i,4] <- X_italy_p_n [i-1,4] + dt * (gamma * X_italy_p_n [i-1,3]) #R
  X_italy_p_n [i,5] <- dt * (alpha * X_italy_p_n [i-1,2]) #new cases
  }

X_italy_p_n <- as.data.frame(X_italy_p_n)
X_italy_p_n$date <- axis_italy_p
  

###########plots#################
#plot: number of new cases main = "Impact of Intervention on the Number of Infectious People Each Day",
plot(X_italy_p[c(1:24), 3] ~ axis_italy_p[c(1:24)] ,  col = 'steelblue4',  lwd = 3,  type = 'l', ylab = "Number of People Infectious", xlab = "Time (Days)",  xlim = as.Date(c("2020-01-31", "2020-04-30")), ylim = c(0,10000))
  lines(X_italy_p[c(24:41), 3] ~ axis_italy_p[c(24:41)] , col = 'violetred4', lwd = 3)
  lines(X_italy_p[c(41:55), 3] ~ axis_italy_p[c(41:55)], col = 'chartreuse4', lwd = 3)
  lines(X_italy_p[c(55:91), 3] ~ axis_italy_p[c(55:91)], col = 'red', lwd = 3)
  lines(X_italy_p_n[c(1:91), 3] ~ axis_italy_p[c(1:91)] , col = 'pink', lwd = 3) #if we are assuming that transmission continued at a high rate
legend("topright", legend = c("R=3.3", "R=2.6", "R=1.5", "R=0.5", "No Intervention"), fill = c("steelblue4", "violetred4", "chartreuse4", "red", "pink"), col = c("steelblue4", "violetred4", "chartreuse4", "red", "pink"), cex = 0.75)



#cumulative cases comparison
#intervention: main = "Comparing the Model Interventions with Non-intervention",
plot(cumsum(X_italy_p[c(1:91), 5]) ~ axis_italy_p[c(1:91)] ,  col = 'navy',  lwd = 3,  type = 'l', ylab = 'Cumulative Cases', xlab = "Time (Days)",  xlim = as.Date(c("2020-01-31", "2020-04-30")), ylim = c(0,50000)) #with intervetnion
 lines(cumsum(X_italy_p_n[c(1:91), 5]) ~ axis_italy_p[c(1:91)] , col = 'red', lwd = 3) 
 legend("topright", legend = c("With Interventions", "Current Situation","Without Interventions"), fill = c("navy", "chartreuse4" ,"red"), col = c("navy", "chartreuse4" ,"red"), cex = 0.5)
 lines(cum.cases.italy[c(11:57), 2] ~ axis_italy_p[c(1:47)], col = 'chartreuse4', lwd = 3) 

 #current italy if we are looking at the cumulative infectious cases that result from the model when the interventions are placed compared to when the interventions are not in place, we see that the interventions decrease gradually. When this is compared to the current cumulative cases in Italy, 
 
```


##Model Application intervention A week Earlier!

I applied interventions put in place to slow transmission looked at how that will change incidences. I compared my values to the cumulative cases, looking at the current situation, my model approximation of the current situation and if the intervention actually creates an impact

```{r}

##############################Intervention###########################################

#assumption: random mixing, person to person transmission, no difference in individual immunity, no deaths
# Parameter values
N_italy <- 60480000
I_italy <- 2 
E_italy <- 20 * I_italy
R_italy <- 0 
S_italy <- N_italy-E_italy-I_italy-R_italy
alpha <- 1/5.6 #incubation period
gamma <- 1/5 #recovery period
duration <-5 #days 
# R0 and beta for different phases based on the impact of intervention to minimize transmission #using infromation from more recent papers. Since R0 = ce/D, these interventions are targeting the ce (contact rate) to reduce transmision. 


R0_1 <- 1.5 #low end (the CI from JHU and authors)
R0_2 <- 2.6 #what was calculated as gthe potential R
R0_3 <- 3.5 #high end (the CI from JHU and authors)
R0_4 <- 0.5 #updated: reduces contact by up to 85% because of lockdown
beta_1 <- R0_1/(N_italy * duration)
beta_2 <- R0_2/(N_italy * duration)
beta_3 <- R0_3/(N_italy * duration)
beta_4 <- R0_4/(N_italy * duration)

#time specification
dt <- 1
start <- 0
end_italy_p <- 90 #days
time_italy_p <- seq(start, end_italy_p, by = dt)
niter_italy_p <- (end_italy_p - start) / dt + 1
s_date_italy_p <- mdy("1-31-2020")
e_date_italy_p <- mdy("4-30-2020")
axis_italy_p <- seq(from = s_date_italy_p, to = e_date_italy_p, by = "day")


X_italy_p_e<- array(data = 0, dim = c(niter_italy_p, 5)) #created an array to catch the results
colnames(X_italy_p_e) <- c("S", "E", "I","R", "New Cases")

X_italy_p_e[1, 1] <- S_italy #initial susuceptible population 
X_italy_p_e[1, 2] <- E_italy #initial latent population 
X_italy_p_e[1, 3] <- I_italy #initial infectious population 
X_italy_p_e[1, 4] <- R_italy #initial recovered population 
X_italy_p_e[1, 5] <- I_italy #new cases

for (i in 2:niter_italy_p) { 
  if(i < 18) {
    #phase 1 transmission with an R0 of 3.1 (January 31 - February 23rd)
    X_italy_p_e[i,1] <- X_italy_p_e[i-1, 1] + dt * (-beta_3 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3] )  #S
    X_italy_p_e[i,2]<- X_italy_p_e[i-1,2] + dt * ((beta_3 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3]) - alpha * X_italy_p_e[i-1,2]) #E
    X_italy_p_e[i,3] <- X_italy_p_e[i-1,3] + dt * (alpha * X_italy_p_e[i-1,2] - gamma * X_italy_p_e[i-1,3])  #I 
    X_italy_p_e[i,4] <- X_italy_p_e[i-1,4] + dt * (gamma * X_italy_p_e[i-1,3]) #R
    X_italy_p_e[i,5] <- dt * (alpha * X_italy_p_e[i-1,2])
  }
  else {
    if ( i >= 18 & i < 34) { 
      #phase 2 transmission with an R0 of 1.9 (February 24 -  March 8) blocking the north
      X_italy_p_e[i,1] <- X_italy_p_e[i-1, 1] + dt * (-beta_2 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3] )  #S
      X_italy_p_e[i,2]<- X_italy_p_e[i-1,2] + dt * ((beta_2 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3]) - alpha * X_italy_p_e[i-1,2]) #E
      X_italy_p_e[i,3] <- X_italy_p_e[i-1,3] + dt * (alpha * X_italy_p_e[i-1,2] - gamma * X_italy_p_e[i-1,3])  #I 
      X_italy_p_e[i,4] <- X_italy_p_e[i-1,4] + dt * (gamma * X_italy_p_e[i-1,3]) #R
      X_italy_p_e[i,5] <- dt * (alpha * X_italy_p_e[i-1,2])
    }
    else{
      if( i >= 34 & i < 48){
        #phase 3 transmission with an R0 of 0.65 (March 8 - March 10) North total Lockdown
        X_italy_p_e[i,1] <- X_italy_p_e[i-1, 1] + dt * (-beta_1 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3] )  #S
        X_italy_p_e[i,2]<- X_italy_p_e[i-1,2] + dt * ((beta_1 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3]) - alpha * X_italy_p_e[i-1,2]) #E
        X_italy_p_e[i,3] <- X_italy_p_e[i-1,3] + dt * (alpha * X_italy_p_e[i-1,2] - gamma * X_italy_p_e[i-1,3])  #I 
        X_italy_p_e[i,4] <- X_italy_p_e[i-1,4] + dt * (gamma * X_italy_p_e[i-1,3]) #R
        X_italy_p_e[i,5] <- dt * (alpha * X_italy_p_e[i-1,2])  
      }
      else{
        X_italy_p_e[i,1] <- X_italy_p_e[i-1, 1] + dt * (-beta_4 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3])  #S
        X_italy_p_e[i,2]<- X_italy_p_e[i-1,2] + dt * ((beta_4 * X_italy_p_e[i-1,1] * X_italy_p_e[i-1,3]) - alpha * X_italy_p_e[i-1,2]) #E
        X_italy_p_e[i,3] <- X_italy_p_e[i-1,3] + dt * (alpha * X_italy_p_e[i-1,2] - gamma * X_italy_p_e[i-1,3])  #I 
        X_italy_p_e[i,4] <- X_italy_p_e[i-1,4] + dt * (gamma * X_italy_p_e[i-1,3]) #R
        X_italy_p_e[i,5] <- dt * (alpha * X_italy_p_e[i-1,2])  
      }
    }
  }
}

X_italy_p_e<- as.data.frame(X_italy_p_e)
X_italy_p_e$date <- axis_italy_p


##############################Comparison with no iuntervention###########################################
#no intervention using high transmission rate 
X_italy_p_e_n <- array(data = 0, dim = c(niter_italy_p, 5)) #created an array to catch the results
colnames(X_italy_p_e_n) <- c("S", "E", "I","R", "New Cases")

X_italy_p_e_n[1, 1] <- S_italy #initial susuceptible population 
X_italy_p_e_n[1, 2] <- E_italy #initial latent population 
X_italy_p_e_n[1, 3] <- I_italy #initial infectious population 
X_italy_p_e_n[1, 4] <- R_italy #initial recovered population 
X_italy_p_e_n[1, 5] <- I_italy #new cases

for(i in 2:niter_italy_p) {
  X_italy_p_e_n [i,1] <- X_italy_p_e_n [i-1, 1] + dt * (-beta_3 * X_italy_p_e_n [i-1,1] * X_italy_p_e_n[i-1,3] )  #S
  X_italy_p_e_n [i,2]<- X_italy_p_e_n [i-1,2] + dt * ((beta_3 * X_italy_p_e_n [i-1,1] * X_italy_p_e_n[i-1,3]) - alpha * X_italy_p_e[i-1,2]) #E
  X_italy_p_e_n [i,3] <- X_italy_p_e_n [i-1,3] + dt * (alpha * X_italy_p_e_n [i-1,2] - gamma * X_italy_p_e_n [i-1,3])  #I 
  X_italy_p_e_n [i,4] <- X_italy_p_e_n [i-1,4] + dt * (gamma * X_italy_p_e_n [i-1,3]) #R
  X_italy_p_e_n [i,5] <- dt * (alpha * X_italy_p_e_n [i-1,2]) #new cases
}

X_italy_p_e_n <- as.data.frame(X_italy_p_e_n)
X_italy_p_e_n$date <- axis_italy_p


###########plots#################
#plot: number of new cases main = "Impact of Intervention on the Number of Infectious People Each Day",
plot(X_italy_p_e[c(1:18), 3] ~ axis_italy_p[c(1:18)] , type = 'l', col = 'steelblue4',  lwd = 3,  ylab = "Number of People Infectious", xlab = "Time (Days)",  xlim = as.Date(c("2020-01-31", "2020-04-30")), ylim = c(0,10000))
lines(X_italy_p_e[c(18:34), 3] ~ axis_italy_p[c(18:34)] , col = 'violetred4', lwd = 3)
lines(X_italy_p_e[c(34:48), 3] ~ axis_italy_p[c(34:48)], col = 'chartreuse4', lwd = 3)
lines(X_italy_p_e[c(48:91), 3] ~ axis_italy_p[c(48:91)], col = 'red', lwd = 3)
lines(X_italy_p[c(24:41), 3] ~ axis_italy_p[c(24:41)] , col = 'violetred4', lwd = 3)
lines(X_italy_p[c(41:55), 3] ~ axis_italy_p[c(41:55)], col = 'chartreuse4', lwd = 3)
lines(X_italy_p[c(55:91), 3] ~ axis_italy_p[c(55:91)], col = 'red', lwd = 3)
lines(X_italy_p_e_n[c(1:91), 3] ~ axis_italy_p[c(1:91)] , col = 'pink', lwd = 3) #if we are assuming that transmission continued at a high rate
lines(X_italy_p[c(1:24), 3] ~ axis_italy_p[c(1:24)] ,  col = 'steelblue4',  lwd = 3)

legend("topright", legend = c("R=3.3", "R=2.6", "R=1.5", "R=0.5", "No Intervention"), fill = c("steelblue4", "violetred4", "chartreuse4", "red", "pink"), col = c("steelblue4", "violetred4", "chartreuse4", "red", "pink"), cex = 0.75)



#cumulative cases comparison
#intervention: main = "Comparing the Model Interventions with Non-intervention",
plot(cumsum(X_italy_p_e[c(1:91), 5]) ~ axis_italy_p[c(1:91)] , type = 'l', col = 'navy',  lwd = 3,  ylab = 'Cumulative Cases', xlab = "Time (Days)",  xlim = as.Date(c("2020-01-31", "2020-04-30")), ylim = c(0,50000)) #with intervetnion
lines(cumsum(X_italy_p_e_n[c(1:91), 5]) ~ axis_italy_p[c(1:91)] , col = 'red', lwd = 3) 
legend("topright", legend = c("With Interventions", "Current Situation","Without Interventions"), fill = c("navy", "chartreuse4" ,"red"), col = c("navy", "chartreuse4" ,"red"), cex = 0.5)
lines(cum.cases.italy[c(11:57), 2] ~ axis_italy_p[c(1:47)], col = 'chartreuse4', lwd = 3) 
lines(cumsum(X_italy_p[c(1:91), 5]) ~ axis_italy_p[c(1:91)] ,  col = 'navy',  lwd = 3)

#current italy if we are looking at the cumulative infectious cases that result from the model when the interventions are placed compared to when the interventions are not in place, we see that the interventions decrease gradually. When this is compared to the current cumulative cases in Italy, 

```
